Multisequence algorithm for coarse-grained biomolecular simulations: application to protein fold switching

### 2017-05-29 Manuscript.tex
TableI and TableII in Manuscript.tex Modified. MCTracFolds.png, network.png, Stderr.pdf, meanE.pdf andSupplementary.tex created. Labels also modified on graphs

A well-known generalized-ensemble algorithm is simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} in which it is the temperature $T$ that becomes a dynamic parameter. In this scheme, frequent visits to high-$T$ allow simulations to readily escape from local energy traps. The ST algorithm thus simulates the joint probability distribution 
\begin{equation}
P(\rbar,m) =\dfrac{1}{\hat{Z}} e^{-\beta_m E(\rbar) + g_m}\,,
\label{ST}
\end{equation}
where  $\beta_m=1/k_\mathrm{B} T_m$, $\{T_m\}_{m=1}^\mathrm{M}$ a set of temperatures and $k_\mathrm{B}$ is Boltzmann's constant. The normalization constant in Eq.~\ref{ST} is  
\begin{equation}
\hat{Z} = \sum_r \sum_{m=1}^{\mathrm{M}}e^{-\beta_m E(\rbar) + g_m}\,,
\end{equation}
where the first sum is over all conformations $\rbar$. The simulation parameters $g_m$ control the marginal probability distribution
\begin{equation}
P(k) = \frac{1}{\hat{Z}}\sum_r e^{-\beta_m E(\rbar) + g_m} \,,
\end{equation}
and must therefore be carefully chosen. A common and convenient choice is $g_m\approx \beta_m F_m$, where $F_m$ is the free energy at temperature $T_m$. With this choice, $P(k)$ becomes approximately flat ensuring all temperatures are frequently visited.

@@

The sequences studied in this work are summarized in Table~I.

\begin{table}
\caption{\label{tab1} List of model sequences of different lengths $N$ studied in this work.}
\begin{ruledtabular}
\begin{tabular}{lcr}
Name & $N$ & Sequence\footnote{In addition, we study two sets of 1,024 sequences derived from the A1-N1 and A2-TN pairs (see text) denoted $\mathrm{S16}_{1024}$  and $\mathrm{S35}_{1024}$, respectively, and a restricted set $\mathrm{S16}_{144}$ with 144 sequences taken from $\mathrm{S16}_{1024}$ ($\mathrm{S16}_{144}$ was also studied in Ref.~\protect\citenum{Holzgrafe2014}). } \\
\hline
A1 & 16 & pphpphhpphpphhpp \\
N1 & 16 & phphphpttphphphp \\
R1 & 16 & pphhphptthpphhpp\\
R2 & 16 & ppphphhtthhphppp\\
A2 & 35 & (A1)ttt(A1)\\
TN & 35 & (A1)ttt(N1)\\
\end{tabular}
\end{ruledtabular}
\end{table}

@@

\begin{table}
\caption{\label{tab2} List of simulations carried out in this work. }
\begin{ruledtabular}
\begin{tabular}{lccccr}
Runs & Algorithm & $\kb T$  & MC steps/run &  Sequences\\
\hline
32 & ST & 0.43--0.65 & $8\times 10^6$ &A1\\ 
32 & ST & 0.43--0.65 & $8\times 10^6$ &N1\\ 
32 & ST & 0.43--0.65 & $8\times 10^6$ &R1\\ 
32 & ST & 0.43--0.65 & $8\times 10^6$ &R2\\ 
32$\times$8\footnote{32  runs per temperature for 8 temperatures in the range given.} & MS &0.43--0.65& $144\times 10^6$ & $\mathrm{S16}_{144}$\\
16 & MS & 0.43  & $5\times 10^9$ &  $\mathrm{S16}_{1024}$ \\
16 & MS & 0.46 & $4\times 10^9$ &  $\mathrm{S35}_{1024}$ \\
\end{tabular}
\end{ruledtabular}
\end{table}

@@

\begin{figure*}
\rotatebox{0}{\includegraphics[width=16.6cm]{ntwk16}}\\
\rotatebox{0}{\includegraphics[width=12.6cm]{ntwk35}}
\caption{Networks of the A1-N1 (top) and A2-TN (bottom) binary sequence space showing the possible stable pathways from one fold to another. }
\end{figure*}

>>>>

A well-known generalized-ensemble algorithm is simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} in which it is the temperature $T$ that becomes a dynamic parameter. In this scheme, frequent visits to high-$T$ allow simulations to readily escape from local energy traps. The ST algorithm thus simulates the joint probability distribution 
\begin{equation}
P(\rbar,m) =\dfrac{1}{\hat{Z}} e^{-\beta_m E(\rbar) + g_m}\,,
\label{ST}
\end{equation}
where  $\beta_m=1/k_\mathrm{B} T_m$, $\{T_m\}_{m=1}^\mathrm{M}$ a set of temperatures and $k_\mathrm{B}$ is Boltzmann's constant. The normalization constant in Eq.~\ref{ST} is  
\begin{equation}
\hat{Z} = \sum_r \sum_{m=1}^{\mathrm{M}}e^{-\beta_m E(\rbar) + g_m}\,,
\end{equation}
where the first sum is over all conformations $\rbar$. The simulation parameters $g_m$ control the marginal probability distribution
\begin{equation}
P(m) = \frac{1}{\hat{Z}}\sum_r e^{-\beta_m E(\rbar) + g_m} \,,
\end{equation}
and must therefore be carefully chosen. A common and convenient choice is $g_m\approx \beta_m F_m$, where $F_m$ is the free energy at temperature $T_m$. With this choice, $P(m)$ becomes approximately flat ensuring all temperatures are frequently visited.

@@

Six model sequences studied in this work are summarized in Table~I. In addition, we study two sets of 1,024 sequences derived from the A1-N1 and A2-TN pairs (see text) denoted $\mathrm{S16}_{1024}$  and $\mathrm{S35}_{1024}$, respectively, and a restricted set $\mathrm{S16}_{144}$ with 144 sequences taken from $\mathrm{S16}_{1024}$ ($\mathrm{S16}_{144}$ was also studied in Ref.~\protect\citenum{Holzgrafe2014}).

\begin{table}
\caption{\label{tab1} List of 6 model sequences of different lengths $N$ studied in this work.}
\begin{ruledtabular}
\begin{tabular}{lcr}
Name & $N$ & Sequence \\
\hline
A1 & 16 & pphpphhpphpphhpp \\
N1 & 16 & phphphpttphphphp \\
R1 & 16 & pphhphptthpphhpp\\
R2 & 16 & ppphphhtthhphppp\\
A2 & 35 & (A1)ttt(A1)\\
TN & 35 & (A1)ttt(N1)\\
\end{tabular}
\end{ruledtabular}
\end{table}

@@

\begin{table}
\caption{\label{tab2} List of simulations carried out in this work. }
\begin{ruledtabular}
\begin{tabular}{lccccr}
Runs & Algorithm & $\kb T$  & MC steps/run\footnote{Excludes thermalization with $10^6$ MC steps.} &  Sequences\\
\hline
32 & ST & 0.43--0.65 & $10^7$ &A1\\ 
32 & ST & 0.43--0.65 & $10^7$ &N1\\ 
32 & ST & 0.43--0.65 & $10^7$ &R1\\ 
32 & ST & 0.43--0.65 & $10^7$ &R2\\ 
32$\times$8\footnote{32  runs per temperature for 8 temperatures in the range given.} & MS &0.43--0.65& $18\times 10^7$ & $\mathrm{S16}_{144}$\\
16 & MS & 0.43  & $5\times 10^9$ &  $\mathrm{S16}_{1024}$ \\
16 & MS & 0.46 & $4\times 10^9$ &  $\mathrm{S35}_{1024}$ \\
\end{tabular}
\end{ruledtabular}
\end{table}

@@

\begin{figure*}
\rotatebox{0}{\includegraphics[width=16.6cm]{network}}
\caption{Networks of the A1-N1 (top) and A2-TN (bottom) binary sequence space showing the possible stable pathways from one fold to another. (light blue) Fold A, (dark red) Fold B, and (black) Bistable. }
\end{figure*}
